---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# reclues
The `reclues` package mimics the SQL Murder Mystery done by [Northwestern's Knight's Lab](https://github.com/NUKnightLab/sql-mysteries) - this uses SQL to solve the mystery. A brief of the task at hand is given in the [walkthrough](http://mystery.knightlab.com/walkthrough.html). 

Want to play the SQL version of the game [online](https://sql-murder-mystery.datasette.io/sql-murder-mystery)?

This R package is set up to import the data as R datasets. 

<!-- badges: start -->
<!-- badges: end -->

The goal of reclues is to provide the datasets and some cheatsheet tidyverse commands for solving the mystery.

# The datasets
The datasets should be available as soon as you install the package. These are the datasets available and the data contained within them.

|  Table Name|  File|  Load with| Note |
|--:|--:|--:|--:|
|  crime_scene_report|  crime_scene_report.rda|  load(file = "crime_scene_report.rda")| date, type, description, city; 1,228 rows|
|  drivers_license|  drivers_license.rda|  load(file = "drivers_license.rda")| id, age, height, eye_color, hair_color, gender, plate_number, car_make, car_model; 10,007 rows|
|  facebook_event_checkin|  facebook_event_checkin.rda|  load(file = "facebook_event_checkin.rda")| person_id, event_id, event_name, date; 20,011 rows|
|  get_fit_now_check_in|  get_fit_now_check_in.rda|  load(file = "get_fit_now_check_in.rda")| membership_id, check_in_date, check_in_time, check_out_time; 2,703 rows|
|  get_fit_now_member|  get_fit_now_member.rda|  load(file = "get_fit_now_member.rda")| id, person_id, name, membership_start_date, membership_status; 184 rows|
|  income|  income.rda|  load(file = "income.rda")| ssn, annual_income; 7,514 rows|
|  interview|  interview.rda|  load(file = "interview.rda")| person_id, transcript; 4,991 rows|
|  person|  person.rda|  load(file = "person.rda")| id, name, license_id, address_number, address_street_name, ssn; 10,011 rows|


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sciencificity/reclues")
```

## Some useful functions to view the data

```{r view, warning=FALSE, message=FALSE}
library(reclues)
library(dplyr)
## basic example code
head(crime_scene_report)
glimpse(crime_scene_report)
```

## Some useful functions to explore the data

### Distinct
Let's say we wanted to see the different __types__ of criminal activity the reports document ....
The `type` field in the crime scene reports table seems to contain this info. We will use the `distinct` function from `dplyr`.

SQL Equivalent is: 

> `SELECT DISTINCT(type) FROM crime_scene_report`


```{r explore, warning=FALSE, message=FALSE}
library(magrittr)

# the magrittr package contains the pipe %>% function
# Take the crime scene report data AND THEN
# give me the distinct values for the `type` variable.
crime_scene_report %>% 
    distinct(type)
```

### Count Distinct
Let's say we were wondering which city has the highest number of crimes ...

SQL Equivalent is:

> SELECT city, count(city) AS n  
> FROM crime_scene_report  
> GROUP BY city  
> ORDER BY n DESC  


```{r explore2}
crime_scene_report %>% 
    count(city) %>% 
    arrange(desc(n))
```
Hhmmm looks like SQL City is quite notorious for crime!

### Like
Let's say we're interested in finding the people that start with a __Z__. We will use the `stringr` package for this. The `str_detect()` function can be used in conjunction with regular expressions - here we looking for names that start with __`(^)`__ __Z__.

SQL Equivalent is:

> SELECT * FROM person  
> WHERE name LIKE 'Z%'  


```{r like, warning=FALSE, message=FALSE}
library(stringr)
person %>% filter(stringr::str_detect(name, "^Z"))
```


